---
title: "week16_recap"
author: "Daniel Trielli"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
library(tidytext)
library(rvest)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")
```

```{r}
# Join the childcare data with the county data 
childcare_inspections_counties <- childcare_inspections |> left_join(md_counties, join_by(fips==GEOID))
```

##### A1:

## The best one to use is NAMELSAD because in the Name column both Baltimore City and Baltimore County are Baltimore.

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
# Calculate the percentage of overdue inspections per county
childcare_inspections_counties <- childcare_inspections_counties |>
group_by(NAMELSAD,overdue,total) |>
summarise(total_county = n()) |>
mutate(pct_over = (overdue/total)*100)
# Create the chart with ggplot
childcare_inspections_counties |>
   ggplot() +
  geom_bar(aes(x=reorder(NAMELSAD, pct_over), weight=pct_over)) +
  coord_flip() +
  labs(
    title="Percent of Overdue Childcare Facility Inspections",
    x = "County",
    y = "Percentage of Overdue Inspections",
    caption = "Source: State of Maryland"
    
  )
```

##### **A2:**

## Montgomery

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
# First, grab data from the ACS. The ACS variable for children age under 5 is B06001_002 We want 2022 data.
md <- get_acs(geography = "county",
              variables = c(under_5 = "B06001_002"),
              state = "MD",
              year = 2022
              )
```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'
md_reformed <- md |>
  rename(kid_pop = estimate) |>
   mutate(NAME = str_replace_all(NAME, ", Maryland", ""))


```

```{r}
childcare_inspections_counties_md <- md_reformed |> left_join(childcare_inspections_counties, join_by(NAME==NAMELSAD))

# Calculate the rates for children per facility
childcare_inspections_counties_md <- childcare_inspections_counties_md |>
mutate(kid_per_fac = (kid_pop/total)) |>
arrange(desc(kid_per_fac))
```

##### **A3:**

## Garrett County

**Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
counties <- counties()
md_counties <- counties |> filter(STATEFP == '24' )

childcare_inspections_counties_md_2 <- md_counties |> inner_join(childcare_inspections_counties_md, join_by(GEOID==GEOID)) 

ggplot() +
  geom_sf(data = childcare_inspections_counties_md_2, aes(fill = kid_per_fac)) +
  theme_minimal() +
  scale_fill_gradient(low = "lightyellow", high = "black")

```

##### **A4**:

Central Maryland and Garrett seem to have a very high ratio.

I used ChatGPT during one portion of my assignment because I couldn't figure out why my map wasn't working and I knew the geometry was in the sheet but R couldn't see it, Turns out I joined my data in the wrong order, and here are the prompts I used to figure it out:

counties \<- counties()
md_counties \<- counties \|\> filter(STATEFP == '24' )\|\>
mutate(GEOID = as.character(GEOID))

childcare_inspections_counties_md_2 \<- childcare_inspections_counties_md \|\> inner_join(md_counties, join_by(GEOID==GEOID))

ggplot() +
geom_sf(data = childcare_inspections_counties_md_2, aes(fill = kid_per_fac)) +
theme_minimal() +
scale_fill_gradient(low = "lightyellow", high = "black") Why isnt my code creating a map

###### ChatGPT said:

There are a few **common issues** in your code that can prevent a map from appearing. Below are the likely causes and how to fix them.

# ✅ **Most Common Problems**

## **1. Your join is not actually joining (empty result)**

If `childcare_inspections_counties_md_2` ends up with **0 rows**, the map will be blank.

### **Typical causes:**

-   

-   

    `GEOID` formats do not match (e.g., `"24001"` vs `24001` vs `"2401"`).

-   

-   

    `inner_join()` removes counties that have no matching inspection data.

-   

### ✔️ Check whether join worked:

```         
nrow(childcare_inspections_counties_md_2) 
```

If you get **0**, that’s why your map is blank.

# **2. Wrong join syntax**

`join_by(GEOID == GEOID)` is ambiguous because both tables have a GEOID column.

### ✔️ Use:

```         
inner_join(md_counties, childcare_inspections_counties_md, by = "GEOID") 
```

Or if using the new `join_by()` syntax:

```         
inner_join(md_counties, childcare_inspections_counties_md, join_by(GEOID == GEOID)) 
```

# **3. Your GEOID coercion may be wrong**

The `{tigris}` `counties()` function already returns `GEOID` as a **character**, so coercing again is unnecessary, but should not hurt.

Verify:

```         
str(md_counties$GEOID) str(childcare_inspections_counties_md$GEOID) 
```

If the inspection dataset has **numeric GEOID**, convert it:

```         
childcare_inspections_counties_md <- childcare_inspections_counties_md |>   mutate(GEOID = sprintf("%05d", as.numeric(GEOID))) 
```

# **4. Variable `kid_per_fac` may be missing**

If the fill column does not exist or contains only `NA`, the plot will draw nothing.

Check:

```         
summary(childcare_inspections_counties_md_2$kid_per_fac) 
```

# ✅ **Corrected Code**

```         
library(tigris) library(dplyr) library(ggplot2)  counties <- counties()  md_counties <- counties |>    filter(STATEFP == "24") |>    mutate(GEOID = as.character(GEOID))  childcare_inspections_counties_md <- childcare_inspections_counties_md |>   mutate(GEOID = as.character(GEOID))  childcare_inspections_counties_md_2 <- md_counties |>   inner_join(childcare_inspections_counties_md, by = "GEOID")  ggplot(childcare_inspections_counties_md_2) +   geom_sf(aes(fill = kid_per_fac)) +   theme_minimal() +   scale_fill_gradient(low = "lightyellow", high = "black") 
```

# If you want, I can help you:

✅ inspect the data\
✅ check your join\
✅ check the GEOID formats\
Just send `dput(head(childcare_inspections_counties_md))` and `dput(head(md_counties))`.

##### You said:

that wasnt helpful

###### ChatGPT said:

No problem — let's get **straight to the reason your map isn't appearing**, based on the code you posted.

# 🎯 **THE REAL PROBLEM**

Your plot:

```         
ggplot() +   geom_sf(data = childcare_inspections_counties_md_2, aes(fill = kid_per_fac)) 
```

**will draw a blank map if the geometry column is missing.**

And **your join likely removed the geometry**.

Why?\
Because you used this:

```         
childcare_inspections_counties_md_2 <- childcare_inspections_counties_md |>    inner_join(md_counties, join_by(GEOID==GEOID)) 
```

This means:

-   

-   

    `childcare_inspections_counties_md` is on the **left**

-   

-   

    `md_counties` (the only dataset containing geometry) is on the **right**

-   

-   

    `inner_join()` keeps the **left-hand column order**, so **geometry is NOT guaranteed to survive unless you join in the other direction**.

-   

**⛔ If the left table is NOT an sf object, the output is also NOT an sf object.**

That means **no geometry → ggplot draws nothing**.

# ✅ **THE FIX**

Join with **the spatial dataset on the left**, so geometry is preserved:

```         
childcare_inspections_counties_md_2 <- md_counties |>   inner_join(childcare_inspections_counties_md, join_by(GEOID == GEOID)) 
```

Or using simpler syntax:

```         
childcare_inspections_counties_md_2 <- md_counties |>   inner_join(childcare_inspections_counties_md, by = "GEOID") 
```

Then plot:

```         
ggplot(childcare_inspections_counties_md_2) +   geom_sf(aes(fill = kid_per_fac)) +   theme_minimal() +   scale_fill_gradient(low = "lightyellow", high = "black") 
```

# ✔️ Why this works

-   

-   

    `md_counties` **is an sf object** → it contains the geometry column.

-   

-   

    Joining **onto** it preserves the geometry.

-   

-   

    Then ggplot knows what to draw.

-   

# If you want, send:

```         
print(childcare_inspections_counties_md_2) 
```

And I can verify if it still has the geometry column (`geom`).

But the **root cause is definitely that your join removed the sf geometry**, making ggplot plot *nothing*.

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.
md_2 <- get_acs(geography = "county",
              variables = c(under_5 = "B19013_001"),
              state = "MD",
              year = 2022
              )
```

```{r}
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'
md_2 <- md_2 |>
rename(median_income = estimate) 
```

```{r}
# Join the dataframes
childcare_median_income <- childcare_inspections_counties_md_2 |> inner_join(md_2, join_by(GEOID==GEOID)) 
```

```{r}
# Create a scatterplot with median income and rate of children per facility
childcare_median_income |>
  ggplot() +
  geom_point(aes(x = median_income, y = kid_per_fac)) +
  labs(
    title="Median Income ",
    x = "Median Income",
    y = "Kids Per Facility",
    caption = "Source: State of Maryland")
```

```{r}
# Create a correlation test.
cor.test(childcare_median_income$median_income, childcare_median_income$kid_per_fac)

```

##### A5: No because the p value is far greater than 0.05

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data
reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()


```

```{r}
# Create a stopword dictionary (feel free to include more words)
data("stop_words")

stop_words <- stop_words |>
  add_row(word = "h") |> 
  add_row(word = "s") |>
  add_row(word = "Compliant") |>
  add_row(word = "Standards") 
 

```

```{r}
# Unnest tokens into bigrams, removing stop words and repasting bigrams, list bigrams
reports_2 <- reports |>
  unnest_tokens(bigram, initial_findings, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  mutate(bigram = paste(word1, word2, sep=" ")) |>
  group_by(bigram) |>
  tally(sort=TRUE) 


```

##### A6: I think that the missing outlets seem to be the most interesting, because I wonder why that is important to note.
